Title here

### 2020-09-04 analysis/paper/paper.tex
borrador

\usepackage{gensymb}
\usepackage{subfig}

@@

\subsubsection{Significance}

@@

Figure \ref{fig:asymsam-timeseries} shows the resulting Asymmetric and
Symmetric time series corresponding to 700 and 50hPa. \#FIXME

At first glance the series can be distinguished by their distributions.
Whereas the tropospheric incides are aproximately normally distributed,
the stratospheric indices are more long-tailed; that is, extreme values
(both negative and positive) abound. The Asymmetric series have both
more variability in the higher frequencies than the Symmetric series.

@@

Figure \ref{fig:trends} shows normalised decadal trends for each index
for the whole period 1979-2918 along with the 95\% confidence interval
in shading for the whole year (row a) and separed by trimesters (rows b
through e). As documented by \#FIXME (e.g. \citet{fogt2020}), there is a
statistically significant increase towards more positive SAM (panel
a.1), which is XX only in Summer and Autumn (panels b.1 and c.1). We
observe these increases mainly in the troposphere, reaching their
maximum at at 100 hPa in Summer. By separating the SAM signal in its
Asymmetric and Symmetric parts, we can not only see that these trends
are almost entirely due to the Symmetric component (columns 2
vs.~columns 3), but in some cases the trends become more clear. In
Summer, the Asymmetric SAM has a statistically non significant negative
trend in the middle troposphere that obscures the signal; as a result,
trends computed using only the Symmetric component are more clear
(compare the shading region in panel b.1 and b.3). In Autumn, using the
Symmetric SAM reveals a statistically significant positive trend in the
stratosphere that is not significant using the Full index.

@@

To undertand the spatial patterns associated with both indeces, we
regressed monthly geopotential anomalies into both indeces using
multiple regression

\begin{figure*}
\includegraphics{2d-regr-1} \caption[Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM]{Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef.}\label{fig:2d-regr}
\end{figure*}

Figure \ref{fig:2d-regr} shows the spatial year-long regression \#FIXME.
Column a are regressinos using the Full SAM, while columns 2 and 3 are
regression coefficients computed in a multiple regression of
geopotential height on the Asymmetric and Symmetric indices at the same
time. Thus, they are to be interpreted as the patterns associated with
each index, controling for the (linear) effect of the other. (Figure A6
\#FIXME illustrates the difference between computing two simple
regressions and one multiple regression.)

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centered in the south pole.
The monopole obtained by multiple regression with the Asymmetric and
Symmetric SAM (panel a.3) is much more symmetric and the shift from
total symmetry is captured by the regression pattern of the Asymmetric
SAM as a wave-1 with maximum anomalies above the Belinghausen Sea on the
Western Hemisphere and and Davids Sea in the Eastern Hemisphere (panel
a.2).

In the troposphere, panel b.1 shows the well known zonally symmetrical
annular mode \emph{contaminated} with zonal asymmetries in the form of a
wave-3. The regression using the Asymmetric and Symmetric SAM indices
successfully disentangle both structures. The Asymmetric component gives
rise to a cleaner zonal wave (panel b.2) and the Symemtric component is
assosiated with an trully annular mode, almost devoid of zonal
asymmetries (panel b.3). note that the wave-3 pattern observed in panel
b.2 is rotated by half a wavelength from the average position of the
mean wave-3 pattern and asociated with \citet{raphael2004}'s ZW3 index
(see Figure 1 from that paper). \#FIXME (agregar algo más?)

@@

The geopotential anomalies associated with the stratospheric SAM (panel
a) are clearly constrained to the stratosphere, which underscores the
disconnect between the stratospheric and tropospheric symetric SAM. The
vertical structure this signal tilts about 60ª to the West between 100
hPa and 1 hPa, suggesting baroclinic processes and polarward transport
of heat (\#FIXME is this ok?). Interestingly, the signal in the
stratosphere maximises near 10 hPa despite using the 50 hPa index for
the regression.

The tropospheric asymmetric SAM has significant signals that extend
upwards the uppermost levels of the reanalysis. In the troposphere, the
wave-3 structure is equivalent barotropic with maximum amplitude at
roughly 250 hPa. \#FIXME moaaaaar.

Interestingly, the structures shown in Figure
\ref{fig:vertical-regression} are surprisignly robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. The pattern cross-correlation
between levels of each segment of the atmosphere is greater than 0.9
(Figure A8). The patterns mainly change in amplitude. The tropospheric
pattern is maximised by the 300 hPa Asymmetric SAM index and the
stratospheric pattern increased monotonically with height.

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the teleconnection pattern associated with the ENSO. Indeed,
\citet{fogt2011} showed that there is a significant relationship between
the SAM and the ENSO. The correlation between the full SAM and the ENSO
as measured by the Multivariate Enso Index \citep{wolter2011} is -0.19.
This relationship is captured entirely the Asymmetric SAM, as this index
has a partial correlation of -0.27 with the MEI, whereas the Symmetric
SAM has null partial correlation with the MEI.

@@

In DJF (column a), the strong negative signal in the tropical Pacific in
panel a.1 is mostly associated with the Asymmetric component (panel
a.2), as is it largely absent in the Symmetric component (panel a.3).
Furthermore, the Asymmetric SAM is also asociated with low temperature
anomalies in the Indian ocean, but this signal is obscured by the
Symmetric variability and thus lost in the Full SAM. Over the
continents, the Asymmetric SAM is assoiated with negative temperature
anomalies which, again, mostly disappear in the Full SAM regression.

The patterns seen in MAM and JJA (columns b and c) are not robustly
significant in the sense that there are no areas with p-values below
0.05 when controlling for FDR following \citet{wilks2016}. Nevertheless,
it is interesting to note that in both trimesters, the sign of the
regression is consistently flipped between the Asymmetric and Symmetric
regressions. In South America, for example, the Asymmetric SAM is
associated with positive temperature anomalies in MAM and negative
temperature anomales in JJA, while the oposite is the case for the
Symemtric SAM.

Finally, in SON (column d), there is no significant temperature signal
associated with the Symmetric SAM (panel d.3), while the Asymmetric SAM
shows a relatively robust signal in the equatorial Pacitic, Australia,
and even Southeast South America. This strong signals are reduced in
intensity in panel a.3.

@@

\begin{figure*}
\includegraphics{pp-regr-oceania-1} \caption[Same but for oceania]{Same but for oceania}\label{fig:pp-regr-oceania}
\end{figure*}

@@

We show that the observed positive trends towards positive SAM is
restricted to the tropospheric SAM and is explained by the Symmetric
component (Figure \textbackslash{}ref\{fig:trends{]}). However, the
degree of asymmetry appears to have increased slightly in the last 40
years, as the Asymmetric SAM explains an increasingly proportion of the
variance (Figure \ref{fig:r-squared-trend}).

In terms of impacts, we see that ..Temperature\ldots{} \#FIXME

The patterns of SAM-associated precipitation anomalies are similarly
well separated. In South America, we show that negative anomalies
observed in Chile related to the SAM are well explained by the Symmetric
compomnet, while the precipitation dipole in Southern South America and
the South Atlantic Convergence Zone is explained by the Asymmetric
component.

\subsubsection{Limitations}

>>>>

\usepackage{gensymb}
\usepackage{subfig}
\usepackage[inline]{showlabels}

@@

\subsubsection{Regressions}

Explain multiple regression interpretation \#FIXME.

@@

Figure \ref{fig:asymsam-timeseries} shows the Asymmetric and Symmetric
time series corresponding to 700 and 50hPa. At first glance the series
can be distinguished by their distributions. Whereas the tropospheric
incides are aproximately normally distributed, the stratospheric indices
are more long-tailed; that is, extreme values (both negative and
positive) abound. The Asymmetric series have both more variability in
the higher frequencies than the Symmetric series.

@@

Figure \ref{fig:trends} shows normalised decadal trends for each index
for the whole period 1979-2918 along with the 95\% confidence interval
in shading for the whole year (row a) and separed by trimesters (rows b
through e). As previously documented (e.g. \citet{fogt2020}), there is a
statistically significant increase towards more positive SAM (panel
a.1), which is evident only in Summer and Autumn (panels b.1 and c.1).
We observe these increases mainly in the troposphere, reaching their
maximum at at 100 hPa in Summer. By separating the SAM signal in its
Asymmetric and Symmetric parts, we can not only see that these trends
are almost entirely due to the Symmetric component (columns 2
vs.~columns 3), but in some cases the trends become more clear. In
Summer, the Asymmetric SAM has a statistically non significant negative
trend in the middle troposphere that obscures the signal; as a result,
trends computed using only the Symmetric component are more clear
(compare the shading region in panel b.1 and b.3). In Autumn, using the
Symmetric SAM reveals a statistically significant positive trend in the
stratosphere that is not significant using the Full index.

@@

\begin{figure*}
\includegraphics{2d-regr-1} \caption[Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM]{Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef. Points marked on panel b.2 are the location of the reference points used by \cite{raphael04} for its Zonal Wave 3 index.}\label{fig:2d-regr}
\end{figure*}

To show if, and to what extent, the Asymmetric and Symmetric SAM
inidices indeed capture the asymmetric and symmetric compoment of the
SAM respectively, we computed the spatial regression of geopotential
height anomalies on these indices and the Full SAM index. Figure
\ref{fig:2d-regr} shows these regressions. Regression coefficients in
column a are computed using the Full SAM. Regression coefficients in
columns b and c are computed using multiple regression using the
Asymmetric and Symmetric indices at the same time. Thus, they are to be
interpreted as the patterns associated with each index, removing the
variability (linearly) explained by the other index.

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centered in the south pole.
The monopole obtained by multiple regression with the Asymmetric and
Symmetric SAM (panel c.1) is much more symmetric and the shift from
total symmetry is captured by the regression pattern of the Asymmetric
SAM as a wave-1 with maximum anomalies above the Belinghausen Sea on the
Western Hemisphere and and Davids Sea in the Eastern Hemisphere (panel
b.1).

In the troposphere, panel a.2 shows the well known zonally symmetrical
annular mode \emph{contaminated} with zonal asymmetries in the form of a
wave-3. The regression using the Asymmetric and Symmetric SAM indices
successfully disentangle both structures. The Asymmetric component gives
rise to a cleaner zonal wave (panel b.2) and the Symemtric component is
assosiated with an trully annular mode, almost devoid of zonal
asymmetries (panel c.2). The wave-3 pattern observed in panel b.2 is
rotated by half a wavelength from the average position of the mean
wave-3 pattern asociated with \citet{raphael2004}'s ZW3 index, whose
reference locations are marked with points in the figure. Thus, the
tropospheric Asymmetric SAM index responds to zonal displacements in the
position of the wave-3 pattern.

@@

The geopotential anomalies associated with the stratospheric SAM (panel
a) are clearly constrained to the stratosphere, which underscores the
disconnect between the stratospheric and tropospheric Asymmetric SAM.
The vertical structure this signal tilts about 60ª to the West between
100 hPa and 1 hPa, suggesting baroclinic processes and polarward
transport of heat (\#FIXME is this ok?). Interestingly, the signal in
the stratosphere maximises near 10 hPa despite using the 50 hPa index
for the regression.

The tropospheric Asymmetric SAM (panel b) has significant signals that
extend upwards to the uppermost levels of the reanalysis. In the
troposphere, the wave-3 structure is equivalent barotropic with maximum
amplitude at roughly 250 hPa. The anomalies are much more intense in the
Western hemisphere, where they extent into the stratosphere. In the
Eastern hemisphere the wave-3 signal is weaker and confined to the
troosphere while negative anomalies dominate in the stratosphere. So,
while the tropospheric Asymmetric SAM index is associated with
stratospheric geopotential anomalies, these do not project strongly onto
the stratospheric Asymmetric SAM.

The structures shown in panels a and b in Figure
\ref{fig:vertical-regression} are surprisignly robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. Pattern cross-correlation is
greater than 0.9 within stratospheric index levels and tropospheric
index levels and less than .4 accross stratospheric and tropospheric
levels (Figure A8), i.e.~regressions obtained using Asymmetric SAM
indices in the troposphere are very similar to each other and different
to the ones obtaines using indices in the stratosphere and vice versa.
The patterns mainly change in amplitude. The tropospheric pattern is
maximised by the 300 hPa Asymmetric SAM index and the stratospheric
pattern increased monotonically with height.

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the teleconnection pattern associated with the ENSO. Indeed,
\citet{fogt2011} showed that there is a significant relationship between
the SAM and the ENSO. The correlation between the full SAM and the ENSO
as measured by the Multivariate Enso Index \citep{wolter2011} is -0.19.
This relationship is captured entirely the Asymmetric SAM, as this index
has a partial correlation of -0.27 with the MEI, whereas the Symmetric
SAM's partial correlation with the MEI is essentially null
(\ensuremath{-7.9\times 10^{-4}}).

@@

In DJF (row 1), the strong negative signal in the tropical Pacific in
panel a.1 is mostly associated with the Asymmetric component (panel
b.1), as is it largely absent in the Symmetric component (panel c.1).
Furthermore, the Asymmetric SAM is also asociated with low temperature
anomalies in the Indian ocean, but this signal is obscured by the
Symmetric variability and thus lost in the Full SAM. Over the
continents, the Asymmetric SAM is assoiated with negative temperature
anomalies which, again, mostly disappear in the Full SAM regression.
Temperature regression with the Symmetric SAM (panel c.1) shows a ring
of negative anomalies around Antarctica surrounded by positive
anomalies. This pattern of anomalies is consistent with thermal wind
balance and the intesification and pole-ward migration of the westerlies
commonly linked to the SAM.

The patterns seen in MAM and JJA (rows 2 and 3) are not robustly
significant in the sense that there are no areas with p-values below
0.05 when controlling for FDR following \citet{wilks2016}. Nevertheless,
it is interesting to note that in both trimesters, the sign of the
regression is consistently flipped between the Asymmetric and Symmetric
regressions. In South America, for example, the Asymmetric SAM is
associated with positive temperature anomalies in MAM and negative
temperature anomales in JJA, while the oposite is the case for the
Symemtric SAM.

Finally, in SON (row 4) there is no significant temperature signal
associated with the Symmetric SAM (panel c.4), while the Asymmetric SAM
shows a relatively robust signal in the equatorial Pacitic, Australia,
and even Southeast South America.

@@

\begin{figure*}
\includegraphics{pp-regr-oceania-1} \caption[Regression pattern of precipiation with Asymmetric and Symmetric SAM]{Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas.}\label{fig:pp-regr-oceania}
\end{figure*}

@@

We show that the observed positive trend towards positive SAM is
restricted to the tropospheric SAM and is explained by the Symmetric
component (Figure \textbackslash{}ref\{fig:trends{]}). However, the
degree of asymmetry appears to have increased slightly in the last 40
years, as the Asymmetric SAM explains an increasingly proportion of the
variance (Figure \ref{fig:r-squared-trend}).

Temperature and precipitation anomalies associated with the SAM respond
to various processes in different locations. The Asymmetric and
Symmetric indices do a reasonable job separating some of them. The
strong sumertime temperature anomalies in the equatorial Pacific are
explained by the Asymmetric component, most likely due to its
relationship with ENSO. The Symmetric SAM, on the other hand, captures
the change in meridional temperature gradient, which is linked to the
SAM-zonal wind relationship by thermal wind balance.

The patterns of SAM-associated precipitation anomalies are similarly
well separated. In South America, we show that negative anomalies
observed in Chile related to the SAM are well explained by the Symmetric
component, while the precipitation dipole in Southern South America and
the South Atlantic Convergence Zone is explained by the Asymmetric
component.

\subsubsection*{Limitations}
\addcontentsline{toc}{subsubsection}{Limitations}
